---
title: "Symbol Entropy estimation Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Symbol Entropy estimation Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
devtools::load_all(".")
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SymbolStats)
```

## Estimate the symbol entropy

We can use the package to estimate the symbol entropy for a sequence of 99 binary values using symbols of length $L=3$.


```{r entropy-example}
str(alldata)
entropy_result <- symbol_entropy(data = sample(c(0,1), 99, replace=TRUE), L = 3, overlap = 1)
```

The result is a list with  1) Table with all symbols or states, 2) symbol entropy and  3) maximum theoretical entropy log(1/2**L).


The symbol entropy is in  `entropy_result$Hs`. All the symbols found are in `entropy_result$table`.


### Application in finance

It is also possible to estimate the symbol entropy to a string of returns. The returns must be binarized, so that when return es positive or zero, the value is 1, otherwise 0. An easy way to binarize is using `dplyr` package:

```{r entropy-example2}
library(dplyr)
library(tidyr)
str(alldata)

# The sp500 prices
sp5 <- alldata %>% select('^GSPC') %>% drop_na

# Compute the SP500 daily returns using the `logvariations_computation` function.
sp5_returns <- logvariations_computation(prices = as.numeric(sp5$`^GSPC`), P=2, overlap=1)

# Binarizing the returns
sp5_binary <- as.data.frame(sp5_returns) %>% mutate(across(everything(), ~case_when(.x >= 0 ~ 1, .x < 0 ~ 0, TRUE ~ NA_real_)))

# Estimating entropy
entropy_result <- symbol_entropy(data = sp5_binary$sp5_returns, L = 3, overlap = 1)
```

All results are displayed in a list. For example, to access the frequency table of the symbols found:

```{r entropy-result1}
entropy_result[[1]]
```

and the estimated symbol entropy in:

```{r entropy-result2}
entropy_result[[2]]
```

